Toward theories of deliberated preferences

### 2020-11-12 simple.tex
Mainly descriptive improv; started getting rid of parametric incompatibility relation

A possible objection to the usefulness of this approach should perhaps be lifted before delving into the formal descriptions of this rough sketch.
It may be argued that because this proposition needs a normative axiom, which cannot be validated empirically, it does not fare better than the alternative propositions existing in the literature. 
%\ac{SEU}, for example, also can be considered to claim to obtain reflective preferences on the basis of normative axioms.
I grant that it is impossible to get completely rid of normative axioms, because prescriptions must be linked to descriptions, and this cannot be done in a complete empirical way.
However, it is good to lighten as much as possible the content of normative axioms, and defer as much content as possible to descriptive claims. The descriptive parts of theories can be confronted to observations, a natural way to lift disagreements that has often proved useful in the history of science.
I postulate that the normative axiom proposed here will be consensual, at least for some non trivial concrete instanciations of the proposed approach, among economists, practitioners, external observers and \acp{DM}. Furthermore, its simplicity and a priori obviousness are strong points in its favor. It has, in some intuitive sense that cannot be described formally, less “content”, in the sense of that word used when economists argue that their theories should not have content about the preferences of the individuals (for fear of being paternalistic or dictatorial). Of course, any decision theory has content in a wider sense of imposing some structure to preferences, otherwise, the theory is void. The question is: are the restrictions imposed of a kind that can be judged acceptable with respect to the considered applications of the theory? Favoring judgments that resist against counter-arguments, as is done here, seems to me to capture exactly what is desired in a prescription context.
%Arguably, the normative axiom proposed here is much weaker than those of a theory such as \ac{SEU}, in the sense of having less substance. (This claim can only be valid at an intuitive level: because \ac{SEU} does not use arguments, its axioms cannot be compared formally to mine.) 
The lack of concrete substance in this normative axiom does not prevent theories of \ac{DJ} to propose substantive, non-trivial claims about reflective preferences. It simply moves the debate from the world of discussions between experts about the validity of the axioms to the world of observations of acceptability of arguments by the persons to which these theories claim to be applicable.
It is also worth noting that the proposed axiom has a specific form that may permit \acp{DM} to reconstruct the reasoning proposed by the theory so as to be able to check that they apply to their situation. Detailed exploration of this aspect is deferred to future work.

@@

The field of formal argumentation theory has developed technical approaches to deal with arguments. Such endeavor starts with an “attack” relation, which indicates which arguments argue against which. Various criteria can then be mathematically defined, in the form of a function selecting subsets of arguments given an attack relation, typically representing sets of arguments that “best resist” their counter-arguments. This field has greatly inspired this proposal, in particular, the notion of attack and of reinstatement. This field however does not view the attack relation as observable, but considers it as given, and its methodology is therefore not empirical in the sense seeked for here. (\Citet{cailloux_formal_2019} elaborates on this difference of objective.) 
\commentYM{ici aussi, je trouve la mention de ces approches maladroite, car elles sont mises en exergue comme si elles étaient des manières de traiter le problème qui t'intéresse, ce qu'elles ne sont pas}

The field of Empirical social choice \citep{gaertner_empirical_2012} investigates individuals’s judgments about fairness: in a typical laboratory experiment, subjects are described some setting requiring to distribute resources among a set of individuals, and are asked to choose a distribution that they consider fair. These interesting approaches permit to study empirically the sense of fairness of real individuals, and compare them to social choice theories. They are distinct from the currently proposed approach as they do not involve arguments; and do not specifically aim to prescribe to an individual.

@@

Define $\ar_0$ as a text arguing that the choice for both $P$ and $D$ should be according to their expected revenue, namely, $\$3.9590$ for $P$ and $\$4$ for $D$. Here is a possible text for $\ar_0$. “Imagine you play this bet 100 times. On average, you would win 99 times and lose once, which means gaining 99 × \$4 = \$396 and losing 1 × \$0.10 = \$0.10, hence a total net gain of \$395.90. Thus, your average net gain for the $P$ bet would be \$3.9590. [Similar text for $D$.] Therefore, we suggest you choose $D$ for a higher expected revenue.”
Define $\ar_1$ as a text arguing that the utility of money should matter, not the revenue, and arguing that a loss hurts more than a gain. 
Omitting many other relevant arguments for now (such as observing that utility should be linear for small amounts of money, for example), define $\allargs = \set{\ar_0, \ar_1}$.

@@

Define $\ileadsto$ as follows. Given $\hist$, $\hist \ileadsto (P ≥ D)$ iff, in a situation of $i$ having been exposed, in order, to the arguments in $\hist$, saying: “Considering these arguments, would you please choose either to play $P$, or play $D$, or declare to be indifferent (in which case one will be picked randomly for you)” leads $i$ to choose $P$ than $D$, or declare indifference. $\hist \ileadsto \prop$ is defined similarly for other elements $\prop \in \allprops$.

@@

We now need an axiom that links the observable reactions of the individuals to their \ac{DJ}.
This will be done through the notion of a decisive argument, that is, intuitively, an argument that is able to convince an individual of the acceptability of a proposition whatever the argumentative path the individual has been submitted to. For formal definitions, some supplementary notations must be defined.

@@

\subsection{Coherence and completeness}
A relation $n \subseteq \allprops × \allprops$ is said to be an incompatibility relation iff $\forall \hist \in \allhist, (\prop, \prop') \in n: \hist \ileadsto \prop ⇒ \hist \nileadsto \prop'$. Two propositions $\prop, \prop' \in \allprops$ are said to be incompatible iff $(\prop, \prop') \in n$.

The following proposition shows that \cref{ax:norm} cannot lead to incompatible propositions being considered deliberated. 
\begin{proposition}[Coherence of acceptability]
	\label{th:protcoh}
	Given any normative theory and incompatibility relation $n$,
	$\forall i \in I, (\prop, \propbar) \in n$:
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \prop] ⇒ [\nexists \ar \in \allargs \suchthat \ar \ileadstost \propbar].
	\end{equation}
\end{proposition}
\commentYM{de même que je trouve très ténu et elliptique le lien entre le protocole dans sa description informelle et dans sa définition formelle, de même je trouve très dérangeant de mettre de terme "protocole" dans le nom de ce genre de proposition. Que l'usage de ce terme ait été choisi par provocation ou non, il me semble indispensable de le discuter}
\begin{proof}
	If for some $\ar \in \allargs$, $\ar \ileadstost \prop$, then by definition of $\ileadstost$, $\forall \ar_1 \in \allargs: (\ar_1, \ar) \ileadsto \prop$, thus by definition of $n$, $\forall \ar_1 \in \allargs: (\ar_1, \ar) \nileadsto \propbar$; thus, for any $\ar' \in \allargs$, $\ar' \ileadstost \propbar$ is false, because it requires that $\forall \ar_1 \in \allargs: (\ar', \ar_1) \ileadsto \propbar$, and thus in particular, that $(\ar', \ar) \ileadsto \propbar$.
\end{proof}

@@

\section{Descriptive theories}
\label{sec:descriptive}
\subsection{Informal presentation}
\commentOCf{Je ne me suis pas encore décidé : descriptive theory or empirical theory ? Pour le moment, j’utilise les deux expressions de façon interchangeable.}
A normative theory defines when a given argument is decisive for a given proposition and a given individual: whenever every sequence of arguments having that argument towards the end of the sequence lead the individual to adhere to the proposition. This, in turn, under condition of accepting its normative validity \commentYM{mais c'est quoi cette "normative validiy" ? de quelles conditions parles-tu ?}, permits in principle to know something about the \ac{DJ} of an individual. But this gives no concrete means to know someone’s \ac{DJ}. The set of finite sequence of arguments is infinite. The definition of decisiveness could admittedly be modified to account for an a priori bound on the maximal length of sequences, but this would not solve the practical problem that this set would be too large to be explored thoroughly 
\commentYM{certain que ce terme existe ?}
\commentOCf{Non, thoroughfully n’existe pas, en effet}. 
Furthermore, only some sequences of arguments can be tested for a given individual: the individual’s patience is limited. A supplementary problem might happen in some situations. Once a sequence has been presented to the individual, this cannot necessarily be undone: she might remember some of the arguments, and they may have a lasting influence on her perspective. Let me call this possibility the hypothesis of memory.
The converse hypothesis amounts to postulating that the individual forgets after some time and use an observation protocol that separates the tests in time sufficiently.
\commentYM{je comprends ce que tu veux dire ici car ça fait un bout de temps qu'on parle de ces question. Je doute cependant qu'un lecteur innocent comprenne un traitre mot}
The hypothesis of memory further reduces the possible sequences that can be tested with a given individual.

Achieving knowledge of someone’s \ac{DJ} in practice is tackled by introducing empirical theories of \ac{DJ}. An empirical theory for a given normative theory defines a subset of propositions of the agenda, that it claims are contained in the \acp{DJ} of individuals concerned by the normative theory. They are called the proposition that the theory supports. \commentYM{dans la mesure où, à ce stade, je ne comprends pas bien ce que contient une "normative theory", la séparation avec l' "empirical theory" n'est pas claire pour moi} 
%The theory also claims that it is able to argue in favor of the proposition it supports in such a way that individuals will be convinced of the quality of these arguments. This is accomplished by three binary relations, part of the definition of empiricial theories: one represents its initial arguments, one plans the possible attacks, and one defines its counter-arguments. An empirical theory also defines a concatenation operation able to dynamically build arguments from fragments of arguments.
%The theory claims that when one of its arguments in support of one of its claims is unconvincing to an individual, it is because of one of the anticipated attacks; and the theory claims to be able to provide counter-arguments to such attacks as they come, that will in turn be either convincing to the individual, or fail to be because of some anticipated attack. The theory is also constrained to have “finite depth”, meaning that it must set an a priori limit to the number of such iterations of attacks and counter-arguments.

@@

Such theories satisfy the two properties that are mandatory to achieve what has been announced, namely, to obtain theories of \ac{DJ} whose power rest on a weak normative axiom and on empirically falsifiable claims. First, the claim of the theory that it argues convincingly holds if and only if no falsification sequence exist (falsification sequences are defined formally here below). Second, granted normative adequacy, if indeed a theory argues convincingly, then it is valid, meaning that the proposition that it claims are \acp{DJ} are indeed \acp{DJ}. These two claims are laid out formally in \cref{th:svalid}.

A second goal of this section is to prove that when two theories disagree about incompatible propositions, one of the theories at least must fail some given sequence of empirical tests, which the proof describes. Importantly, these tests can be performed on a single individual and are robust against the hypothesis of memory.

\subsection{Definition}
In this section, a normative theory of \ac{DJ} $(\allprops, \allargs, I, {\allleadsto})$ is supposed given.
A descriptive tuple for a normative theory of \ac{DJ} is defined as a tuple $(\gleadsto, \gbeats, \alldargs, \dbeats, +)$, where ${\gleadsto} \subseteq \allargs × \allprops$ is an initial argumentation relation; ${\gbeats} \subseteq \allargs × \allargs$ is an anticipated attacks relation; $\alldargs$ is a primitive set of argument fragments; ${\dbeats} \subseteq \alldargs × \gbeatsinv(\allargs)$ is a defense relation; and $+$ combines arguments to coalitions of fragments in order to form new arguments.
A descriptive theory for a normative theory of \ac{DJ} is a descriptive tuple that has a bounded depth.
These concepts are now defined formally.
\commentYM{puisque la théorie empirique est une théorie empirique D'UNE CERTAINE THEORIE NORMATIVE, je m'attendrais  à ce qu'on voit apparaître la théorie normative dans la formule de la théorie empirique, non ?}

@@

A defense relation ${\dbeats} \subseteq \alldargs × \gbeatsinv(\allargs)$ is an injective (left-unique) and surjective (right-total) binary relation over $\alldargs × \gbeatsinv(\allargs)$. \commentYM{le fait qu'elle soit surjective et injective est une exigence que tu poses dans la définiton ? pourquoi ? à quoi est-ce que cela sert ?} 
These constraints guarantee that every anticipated attack has exactly one fragment defending against it, thus, $\forall \ar_1 \suchthat [\exists \ar_0 \suchthat \ar_1 \gbeats \ar_0]$: $\dbeatsinv(\ar_1)$ is a singleton. 
Note that $\dbeats$ need not be a bijection as it is generally not a function (it is not restrictive to suppose it left-total, but it generally is not right-unique). 
Given $\ar_1 \in \gbeatsinv(\allargs)$, define $\dbeatsinv(\ar_1)$ as the fragment defending against the anticipated attack $\ar_1$.
\commentYM{dans toute cette sous-section la lecture est rendue très difficile par ce genre de formulation qui assigne une intentionnalité à la théorie. Je ne comprends pas le sens métaphorique de ces phrase, donc je ne les comprends pas}

@@

\subsection{Validity}
\label{sec:validity}
Given $\args \subseteq \allargs$ and $\hist \in \allhist$,
the set $\args \cap \range(\hist)$ denotes the arguments that are both in $\args$ and contained in the sequence $\hist$.

Define a falsification sequence for $i, \ar_0, \phi$, with $\ar_0 \gleadsto \phi$, thus $\ar_0$ an initial argument for $\phi$, as a finite sequence $\hist^1, \hist^3, …, \hist^{N + 1}$, for some $N \in 2\N$, such that it presents successive challenges, which means the following. 
Given $k \in 2\N \suchthat k ≤ N$, $\ar_k \in \allargs$ and $\hist^{k + 1} \in \allhist$, 
define $\args_{k + 1} = \gbeatsinv(\ar_k) \cap \range(\hist^{k + 1})$; 
define $\dargs[k + 2] = \dbeatsinv(\args_{k + 1})$ as those fragments defending against the anticipated attackers $\args_{k + 1}$; 
and, if $\args_{k + 1} ≠ \emptyset$ (equivalently, if $\dargs[k + 2] ≠ \emptyset$), define $\ar_{k + 2} = \ar_k + \dargs[k + 2]$. 
The sequence presents successive challenges iff, $\forall k \in 2\N \suchthat k ≤ N$, we have that $\ar_k \in \histend[k + 1]$, $\hist^{k + 1} \nileadstost \phi$, and $[\args_{k + 1} = \emptyset ⇔ k = N]$.

\begin{definition}[Validity]
	An empirical theory is \emph{valid} iff $\forall i \in I, \prop \in \gpropse: \prop \in \iprops$.
\end{definition}

@@

\subsection{Effectively comparing models}
Given $\prop \in \allprops$ and two empirical theories $\gamma$ and $\delta$ for a given normative theory with an incompatibility relation $n$, say that they disagree about $(\prop, \propbar) \in n$ iff $\prop \in \gpropse[\gamma] \land \propbar \in \gpropse[\delta]$. Say that they are incompatible iff there is at least one pair $(\prop, \propbar) \in \allprops$ that they disagree about.

>>>>

A possible objection to the usefulness of this approach should perhaps be lifted before delving into the formal descriptions of this rough sketch.
It may be argued that because this proposition needs a normative axiom, which cannot be validated empirically, it does not fare better than the alternative propositions existing in the literature. 
%\ac{SEU}, for example, also can be considered to claim to obtain reflective preferences on the basis of normative axioms.
I grant that it is impossible to get completely rid of normative axioms, because prescriptions must be linked to descriptions, and this cannot be done in a complete empirical way.
However, it is good to lighten as much as possible the content of normative axioms, and defer as much content as possible to descriptive claims. The descriptive parts of theories can be confronted to observations, a natural way to lift disagreements that has often proved useful in the history of science.
I postulate that the normative axiom proposed here will be consensual, at least for some non trivial concrete instanciations of the proposed approach, among economists, practitioners, external observers and \acp{DM}. Furthermore, its simplicity and a priori obviousness are strong points in its favor. It has, in some intuitive sense that cannot be described formally, less “content”, in the sense of that word used when economists argue that their theories should not have content about the preferences of the individuals (for fear of being paternalistic or dictatorial). Of course, any decision theory has content in a wider sense of imposing some structure to preferences, otherwise, the theory is void. The question is: are the restrictions imposed of a kind that can be judged acceptable with respect to the considered applications of the theory? Favoring judgments that resist counter-arguments, as done here, seems to me to capture exactly what is desired in a prescription context.
%Arguably, the normative axiom proposed here is much weaker than those of a theory such as \ac{SEU}, in the sense of having less substance. (This claim can only be valid at an intuitive level: because \ac{SEU} does not use arguments, its axioms cannot be compared formally to mine.) 
The lack of concrete substance in this normative axiom does not prevent theories of \ac{DJ} to propose substantive, non-trivial claims about reflective preferences. It simply moves the debate from the world of discussions between experts about the validity of the axioms to the world of observations of acceptability of arguments by the persons to which these theories claim to be applicable.
It is also worth noting that the proposed axiom has a specific form that may permit \acp{DM} to reconstruct the reasoning proposed by the theory so as to be able to check that they apply to their situation. Detailed exploration of this aspect is deferred to future work.

@@

The field of formal argumentation theory has developed technical approaches to deal with arguments. Such endeavor starts with an “attack” relation, which indicates which arguments argue against which. Various criteria can then be mathematically defined, in the form of a function selecting subsets of arguments given an attack relation, typically representing sets of arguments that “best resist” their counter-arguments. This field has greatly inspired this proposal, in particular, the notion of attack and of reinstatement. This field however does not view the attack relation as observable, but considers it as given, and its methodology is therefore not empirical in the sense seeked here. (\Citet{cailloux_formal_2019} elaborates somewhat on this difference of objective.) 
\commentYM{ici aussi, je trouve la mention de ces approches maladroite, car elles sont mises en exergue comme si elles étaient des manières de traiter le problème qui t'intéresse, ce qu'elles ne sont pas}

The field of Empirical social choice \citep{gaertner_empirical_2012} investigates individuals’s judgments about fairness: in a typical laboratory experiment, subjects are described some setting requiring to distribute resources among a set of individuals, and are asked to choose a distribution that they consider fair. These interesting approaches permit to study empirically the sense of fairness of real individuals, and compare them to social choice theories. They are distinct from the currently proposed approach as they seek to obtain their a priori sense of fairness, rather than aim at observing how subjects would appreciate (possibly counter-intuitive) claims about justice when they are backed by (possibly elaborate) arguments; and do not specifically aim to prescribe to an individual.

@@

Define $\ar_0$ as a text arguing that the choice between $P$ and $D$ should be according to their expected revenue, namely, $\$3.959$ for $P$ and $\$4$ for $D$. Here is a possible text for $\ar_0$. “Imagine you play this bet 100 times. On average, you would win 99 times and lose once, which means gaining 99 × \$4 = \$396 and losing 1 × \$0.10 = \$0.10, hence a total net gain of \$395.90. Thus, your average net gain for the $P$ bet would be \$3.9590. [Similar text for $D$.] Therefore, we suggest you choose $D$ for a higher expected revenue.”
Define $\ar_1$ as a text arguing that the utility of money should matter, not the revenue, and arguing that a loss hurts more than a gain. 
Omitting many other relevant arguments for now (such as arguing that utility should be linear for small amounts of money, for example), define $\allargs = \set{\ar_0, \ar_1}$.

@@

Define $\ileadsto$ as follows. Given an argumentative path $\hist$, $\hist \ileadsto (P ≥ D)$ iff, in a situation of $i$ having been exposed, in order, to the arguments in $\hist$, letting $i$ choose between $P$ and $D$ or declare indifference leads $i$ to choose $P$, or declare indifference. Here is a possible text presented to $i$ to ask for a choice: “Considering these arguments, would you please choose either to play $P$, or play $D$, or declare to be indifferent (in which case one will be picked randomly for you)”. 
Similarly, define $\hist \ileadsto (P > D)$ iff showing the arguments in $\hist$ to $i$ and letting $i$ choose between $P$ and $D$ or declare indifference leads $i$ to choose $P$ and not $D$. 
Define similarly $\hist \ileadsto \prop$ for $\prop \in \set{D ≥ P, D > P}$.

@@

We now need an axiom that links the observable reactions of the individuals to their \ac{DJ}.
This will be done through the notion of a decisive argument, that is, intuitively, an argument that is able to convince an individual of the acceptability of a proposition whatever the argumentative path submitted to the individual. For formal definitions, some supplementary notations must be defined.

@@

\subsection{Coherence and completeness}
%A relation $n \subseteq \allprops × \allprops$ is said to be an incompatibility relation iff $\forall \hist \in \allhist, (\prop, \prop') \in n: \hist \ileadsto \prop ⇒ \hist \nileadsto \prop'$. 
%Two propositions $\prop, \prop' \in \allprops$ are said to be incompatible iff $(\prop, \prop') \in n$.
Two propositions $\prop, \prop' \in \allprops$ are said to be incompatible iff $\forall \hist \in \allhist: \hist \ileadsto \prop ⇒ \hist \nileadsto \prop'$. Let $\incompat \subseteq \Phi × \Phi$ denote the set of incompatible pairs of propositions.

The following proposition shows that \cref{ax:norm} cannot lead to incompatible propositions being considered deliberated. 
\begin{proposition}[Coherence of acceptability]
	\label{th:protcoh}
	Given any normative theory,
	$\forall i \in I, (\prop, \propbar) \in \incompat$:
	\begin{equation}
		[\exists \ar \in \allargs \suchthat \ar \ileadstost \prop] ⇒ [\nexists \ar \in \allargs \suchthat \ar \ileadstost \propbar].
	\end{equation}
\end{proposition}
\commentYM{de même que je trouve très ténu et elliptique le lien entre le protocole dans sa description informelle et dans sa définition formelle, de même je trouve très dérangeant de mettre de terme "protocole" dans le nom de ce genre de proposition. Que l'usage de ce terme ait été choisi par provocation ou non, il me semble indispensable de le discuter}
\begin{proof}
	If for some $\ar \in \allargs$, $\ar \ileadstost \prop$, then by definition of $\ileadstost$, $\forall \ar_1 \in \allargs: (\ar_1, \ar) \ileadsto \prop$, thus by definition of incompatibility, $\forall \ar_1 \in \allargs: (\ar_1, \ar) \nileadsto \propbar$; thus, for any $\ar' \in \allargs$, $\ar' \ileadstost \propbar$ is false, because it requires that $\forall \ar_1 \in \allargs: (\ar', \ar_1) \ileadsto \propbar$, and thus in particular, that $(\ar', \ar) \ileadsto \propbar$.
\end{proof}

@@

\section{Descriptive theories}
\label{sec:descriptive}
\subsection{Informal presentation}
\commentOCf{Je ne me suis pas encore décidé : descriptive theory or empirical theory ? Pour le moment, j’utilise les deux expressions de façon interchangeable.}
A normative theory defines when a given argument is decisive for a given proposition and a given individual: whenever every sequence of arguments having that argument towards the end of the sequence lead the individual to adhere to the proposition. This, in turn, under condition of accepting its normative validity \commentYM{mais c'est quoi cette "normative validiy" ? de quelles conditions parles-tu ?}, permits in principle to know something about the \ac{DJ} of an individual. But this gives no concrete means to know anyone’s \ac{DJ}. The set of finite sequence of arguments, thus of argumentative paths, is infinite. The definition of decisiveness could admittedly be modified to account for an a priori bound on the maximal length of sequences, but this would not solve the practical problem that this set would be too large to be explored thoroughly 
\commentYM{certain que ce terme existe ?}
\commentOCf{Non, thoroughfully n’existe pas, en effet}. 
Furthermore, only some argumentative paths can be tested for a given individual: the individual’s patience is limited. A supplementary problem might happen in some situations. Once an argumentative path has been presented to the individual, this cannot necessarily be undone: she might remember some of the arguments, and they may have a lasting influence on her perspective. Let me call this possibility the hypothesis of memory.
The converse hypothesis amounts to postulating that the individual forgets after some time and use an observation protocol that separates the tests in time sufficiently.
\commentYM{je comprends ce que tu veux dire ici car ça fait un bout de temps qu'on parle de ces question. Je doute cependant qu'un lecteur innocent comprenne un traitre mot}
The hypothesis of memory further reduces the possible paths that can be tested with a given individual.

Importantly, theories of \ac{DJ} must also be able to argue differently for different individuals. Even considering individuals with the same \ac{DJ}, different individuals might consider as decisive different arguments. Accepting non trivial propositions often requires to accept several sub-propositions, that together permit to be convinced of the acceptability of the proposition. Depending on which of these sub-propositions the individual already considers to be acceptable, parts of the argumentation can be omitted, and, perhaps, must be omitted: an overly long argument, stating many things that the listener already knows, has fewer chance to convince the listener than a short argument that goes straight to the point about which the listener doubts. 
%Mathematicians do not write their proofs starting from the axioms of set theory: they use results that they assume the reader already is aware of, and are ready (in principle) to supply the proofs of these intermediate results if a reader has doubt about some of them. 
In other words, what is needed for a theory of \ac{DJ} is not just a capacity to produce a unique decisive argument, with the claim that this argument will convince individuals in $I$, but rather a capacity to produce dynamically arguments that depend on which counter-arguments the individual finds appealing, with a claim that this dynamic argumentation strategy will succeed in convincing individuals in $I$.

Achieving practical knowledge of \acp{DJ} is tackled by introducing descriptive theories of \ac{DJ}. They are defined in \cref{sec:descrdef}. An empirical theory for a given normative theory defines a subset of propositions of the agenda, that it claims are contained in the \acp{DJ} of individuals concerned by the normative theory. They are called the proposition that the theory supports. \commentYM{dans la mesure où, à ce stade, je ne comprends pas bien ce que contient une "normative theory", la séparation avec l' "empirical theory" n'est pas claire pour moi} 
%The theory also claims that it is able to argue in favor of the proposition it supports in such a way that individuals will be convinced of the quality of these arguments. This is accomplished by three binary relations, part of the definition of empiricial theories: one represents its initial arguments, one plans the possible attacks, and one defines its counter-arguments. An empirical theory also defines a concatenation operation able to dynamically build arguments from fragments of arguments.
%The theory claims that when one of its arguments in support of one of its claims is unconvincing to an individual, it is because of one of the anticipated attacks; and the theory claims to be able to provide counter-arguments to such attacks as they come, that will in turn be either convincing to the individual, or fail to be because of some anticipated attack. The theory is also constrained to have “finite depth”, meaning that it must set an a priori limit to the number of such iterations of attacks and counter-arguments.

@@

Such theories satisfy the two properties that are mandatory to achieve what has been announced, namely, to obtain theories of \ac{DJ} whose power rest on a weak normative axiom and on empirically falsifiable claims. First, the claim of the theory that it argues convincingly holds if and only if no falsification sequence exist (falsification sequences are defined formally here below). Second, granted normative adequacy, if indeed a theory argues convincingly, then it is valid, meaning that the proposition that it claims are \acp{DJ} are indeed \acp{DJ}. These two claims are laid out formally in \cref{sec:validity}.

This permits to ensure that, under Normative adequacy, Validity is an empirical property. But it does not give a concrete way of arbitrating between conflicting theories. A supplementary important result, described in \cref{sec:compare}, is that when two theories disagree about incompatible propositions, one of the theories at least must fail some given sequence of empirical tests, which the proof describes. These tests can be performed on a single individual and are robust against the hypothesis of memory.

\subsection{Definition}
\label{sec:descrdef}
In this section, a normative theory of \ac{DJ} $(\allprops, \allargs, I, {\allleadsto})$ is supposed given.
A descriptive tuple for a normative theory of \ac{DJ} is defined as a tuple $(\gleadsto, \gbeats, \alldargs, \dbeats, +)$, where ${\gleadsto} \subseteq \allargs × \allprops$ is an initial argumentation relation; ${\gbeats} \subseteq \allargs × \allargs$ is an anticipated attacks relation; $\alldargs$ is a primitive set of argument fragments; ${\dbeats} \subseteq \alldargs × \gbeatsinv(\allargs)$ is a defense relation; and $+$ combines arguments to coalitions of fragments in order to form new arguments.
A descriptive theory for a normative theory of \ac{DJ} is a descriptive tuple that has a bounded depth.
These concepts are now defined formally.
\commentYM{puisque la théorie empirique est une théorie empirique D'UNE CERTAINE THEORIE NORMATIVE, je m'attendrais  à ce qu'on voit apparaître la théorie normative dans la formule de la théorie empirique, non ?}

@@

A defense relation ${\dbeats} \subseteq \alldargs × \gbeatsinv(\allargs)$ is an injective (left-unique) and surjective (right-total) binary relation over $\alldargs × \gbeatsinv(\allargs)$. \commentYM{le fait qu'elle soit surjective et injective est une exigence que tu poses dans la définiton ? pourquoi ? à quoi est-ce que cela sert ?} 
These constraints guarantee that every anticipated attack has exactly one fragment defending against it, thus, $\forall \ar_1 \suchthat [\exists \ar_0 \suchthat \ar_1 \gbeats \ar_0]$, $\dbeatsinv(\ar_1)$ is a singleton. 
Note that $\dbeats$ need not be a bijection as it is generally not a function (it is not restrictive to suppose it left-total, but it generally is not right-unique). 
Given $\ar_1 \in \gbeatsinv(\allargs)$, define $\dbeatsinv(\ar_1)$ as the fragment defending against the anticipated attack $\ar_1$.
\commentYM{dans toute cette sous-section la lecture est rendue très difficile par ce genre de formulation qui assigne une intentionnalité à la théorie. Je ne comprends pas le sens métaphorique de ces phrase, donc je ne les comprends pas}

@@

\subsection{Validity and falsification sequences}
\label{sec:validity}
\begin{definition}[Validity]
	An empirical theory is \emph{valid} iff $\forall i \in I, \prop \in \gpropse: \prop \in \iprops$.
\end{definition}

Given $\args \subseteq \allargs$ and $\hist \in \allhist$,
recall that the set $\args \cap \range(\hist)$ denotes the arguments that are both in $\args$ and contained in the sequence $\hist$.

Define a falsification sequence for $(i, \ar_0, \phi)$, with $\ar_0 \gleadsto \phi$, thus $\ar_0$ an initial argument for $\phi$, as a finite sequence $\hist^1, \hist^3, …, \hist^{N + 1}$, for some $N \in 2\N$, such that it presents successive challenges, which means the following. 
Given $k \in 2\N \suchthat k ≤ N$, $\ar_k \in \allargs$ and $\hist^{k + 1} \in \allhist$, 
define $\args_{k + 1} = \gbeatsinv(\ar_k) \cap \range(\hist^{k + 1})$; 
define $\dargs[k + 2] = \dbeatsinv(\args_{k + 1})$ as those fragments defending against the anticipated attackers $\args_{k + 1}$; 
and, if $\args_{k + 1} ≠ \emptyset$ (equivalently, if $\dargs[k + 2] ≠ \emptyset$), define $\ar_{k + 2} = \ar_k + \dargs[k + 2]$. 
The sequence presents successive challenges iff, $\forall k \in 2\N \suchthat k ≤ N$, we have that $\ar_k \in \histend[k + 1]$, $\hist^{k + 1} \nileadstost \phi$, and $[\args_{k + 1} = \emptyset ⇔ k = N]$.

A descriptive theory has no falsification sequence iff $\forall (i, \ar_0, \phi)$, no finite sequence of paths present successive challenges.

@@

\subsection{Effectively comparing models}
\label{sec:compare}
Given $\prop \in \allprops$ and two empirical theories $\gamma$ and $\delta$ for a given normative theory with an incompatibility relation $n$, say that they disagree about $(\prop, \propbar) \in n$ iff $\prop \in \gpropse[\gamma] \land \propbar \in \gpropse[\delta]$. Say that they are incompatible iff there is at least one pair $(\prop, \propbar) \in \allprops$ that they disagree about.
